---
title: 'Reproducible Research: Peer Assessment 1'
output: html_document
---
Loading and preprocessing the data 

What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

1. Make a histogram of the total number of steps taken each day

2. Calculate and report the mean and median total number of steps taken per day
```{r}
library(knitr)
opts_chunk$set(echo = TRUE)

library(dplyr)
library(lubridate)
library(ggplot2)

data <- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character","integer"))

data$date <- ymd(data$date)

steps <- data %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```


A histogram of the total number of steps taken each day:

```{r, echo=FALSE}
ggplot(steps, aes(x = steps)) + geom_histogram(fill = "firebrick", binwidth = 1000) + labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")
```

```{r}
mean_steps <- mean(steps$steps, na.rm = TRUE)
median_steps <- median(steps$steps, na.rm = TRUE)
mean_steps
median_steps
```
What is the average daily activity pattern?
```{r}
interval <- data %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(steps = mean(steps))
```
The time series of the 5-minute interval and average steps taken:

```{r, echo=FALSE}
ggplot(interval, aes(x=interval, y=steps)) + geom_line(color = "firebrick")
```
Use which.max() to find out the maximum steps, on average, across all the days

```{r}
interval[which.max(interval$steps),]
```
Imputing missing values

```{r}
sum(is.na(data$steps))

data_full <- data
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]
```

Calculate the number of steps taken in each 5-minute interval per day using dplyr and group by interval. Use ggplot for making the histogram:
```{r}
steps_full <- data_full %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```
Create graph:
```{r, echo=FALSE}
ggplot(steps_full, aes(x = steps)) + geom_histogram(fill = "firebrick", binwidth = 1000) + labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")
```
Calculate the mean and median steps with the filled in values:
```{r}
mean_steps_full <- mean(steps_full$steps, na.rm = TRUE)
median_steps_full <- median(steps_full$steps, na.rm = TRUE)
mean_steps_full
median_steps_full
```
Are there differences in activity patterns between weekdays and weekends?
Use dplyr and mutate to create a new column, weektype, and apply whether the day is weekend or weekday:
```{r}
data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)
```
Calculate the average steps in the 5-minute interval and use ggplot for making the time series of the 5-minute interval for weekday and weekend, and compare the average steps:
```{r}
interval_full <- data_full %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
s <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) + geom_line() + facet_wrap(~weektype, ncol = 1, nrow=2)
print(s)
```
